---
title: "BioGPS expression analysis"
author: "John Lee"
date: "2017å¹?8æœ?14æ—?"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load the packages
```{r loading, message=FALSE}
if (!require(Biobase)) {
    source("http://bioconductor.org/biocLite.R")
    biocLite("Biobase")
    library(Biobase)
}
if (!require(GEOquery)) {
    source("http://bioconductor.org/biocLite.R")
    biocLite("GEOquery")
    library(GEOquery)
}
if (!require(ggplot2)) {
    install.packages("ggplot2")
    library(ggplot2)
}
if (!require(tidyr)) {
    install.packages("tidyr")
    library(tidyr)
}
setwd("E:\\Biological_DATA\\BioGPS")
```
## read the data and choose genes needed.
Here the geo acession number and the platform is required. We used the Mouse MOE430 Gene Atlas as an example of how to analyze the expression of a specific gene in different types of cells.
We get the expression data and the experiment design `description`, and we get 
the plantform information in `platf_table`.

```{r}
if (file.exists("E:\\Biological_DATA\\BioGPS\\GSE10246.RData")){
    load("E:\\Biological_DATA\\BioGPS\\GSE10246.RData")
} else {
    gset <- getGEO("GSE10246", GSEMatrix =TRUE, getGPL=FALSE)
    # This returns a list of gsets, usually the length of the gset is 1
    if (length(gset) > 1) idx <- grep("GPL1261", attr(gset, "names")) else idx <- 1
    gset <- gset[[idx]]
    # get the first gset
    ex <- exprs(gset)
    # expression data
    descriptions <- gset@phenoData@data$description
    # descriptions,here it is different types of cells
    platf <-getGEO(annotation(gset), AnnotGPL=TRUE,destdir="E:\\Biological_DATA\\BioGPS")
    platf_table <-data.frame(attr(dataTable(platf), "table"))
    save(ex,gset,descriptions, platf_table, file =
             "E:\\Biological_DATA\\BioGPS\\GSE10246.RData")
}
```

## Getting the desired gene expression data.
The major problem here is that one gene may contain several probes. In bioGPS
website they only used the most active probe. We can also do the same trick, but
checking the consistency is always prefered.
```{r}
genes <- c("Foxo1","Foxo3","Nfe2l2")
for (gene in genes) {
    print(gene)
    ids <- which(platf_table$Gene.symbol==gene)
    for (id in ids) {
        print(rownames(ex)[id])
        print(summary(ex[id,]))
    }
    print(cor(t(ex[ids,])))
}
cells <- data.frame(descriptions, colId=1:length(descriptions))
```
## Choose probes and cells to plot the final data
In many cases, the first probe is the best one with high expression and high correlation
with other probes. So our probes are Foxo1: 1416981_at, Foxo3: 1434831_a_at,
Nfe2l2: 1416543_at

The cells we are choosing is:
1. T-cells
2. B-cells
3. macrophages
4. dendritic cells
5. Thymocyte
```{r}
cellselect <- cells[grep("B-cells",cells$descriptions),]
cellselect <- rbind(cellselect,
              cells[grep("T-cells|macrophage|dendritic|thymocyte",cells$descriptions),])
cellselect$descriptions <- as.character(cellselect$descriptions)
cellselect[3:4,1] <-"B-cells_follicular"

exselect <- t(ex[match(c("1416981_at","1434831_a_at","1416543_at"),rownames(ex)),
               cellselect[,2]])
plotData <- data.frame(exselect, as.character(cellselect$descriptions))
names(plotData) <- c("Foxo1","Foxo3","Nfe2l2","celltype")
plotData <- gather(plotData,gene,exp,-celltype)
plotData[73:108,3] <- plotData[73:108,3]/4
# get the mean and the sd
myData <- aggregate(plotData$exp,
    by = list(plotData$celltype, plotData$gene),
    FUN = function(x) c(mean = mean(x), sd = sd(x),
                        n = length(x)))
myData <- do.call(data.frame, myData)
names(myData) <- c("celltype", "gene", "mean", "sd", "n")
```
## Final plot procedure.
```{r}
pdf(file = "Expression of Foxo1, Foxo3 and Nfe2l2 in different immune cells.pdf",
    width=12,height = 7)
limits <- aes(ymax = myData$mean + myData$sd,
              ymin = myData$mean - myData$sd)
p <- ggplot(data = myData, aes(x = gene, y = mean,
               fill = celltype))
p + geom_bar(stat = "identity",position = position_dodge(0.9))+
  geom_errorbar(limits, position = position_dodge(0.9), width = 0.25) +
  labs(x = "cell types", y = "relative expression") +
  ggtitle("Expression of Foxo1, Foxo3 and Nfe2l2 in different immune cells") +
  scale_fill_discrete(name = "genes") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
dev.off()
```
Finally the plot need to be saved, you can also plot it.