---
title: "GSE analysis of GEO data"
author: "John Li"
date: "July 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load R libraries required
version info:

```{r load library}
if (!require(Biobase)) {
    source("http://bioconductor.org/biocLite.R")
    biocLite("Biobase")
    library(Biobase)
}
if (!require(GEOquery)) {
    source("http://bioconductor.org/biocLite.R")
    biocLite("GEOquery")
    library(GEOquery)
}
if (!require(limma)) {
    source("http://bioconductor.org/biocLite.R")
    biocLite("limma")
    library(limma)
}
if (!require(apcluster)) {
    source("http://bioconductor.org/biocLite.R")
    biocLite("apcluster")
    library(apcluster)
}
if (!require(affycoretools)) {
    source("http://bioconductor.org/biocLite.R")
    biocLite("affycoretools")
    library(affycoretools)
}
if (!require(ggplot2)) {
    install.packages("ggplot2")
    library(ggplot2)
}
setwd("E:/biological_DATA/GEO_data/")
```

## 
Here GSE name and platform is required.

```{r }
gset <- getGEO("GSE4528", GSEMatrix =TRUE, AnnotGPL=TRUE)
if (length(gset) > 1) idx <- grep("GPL339", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]

# make proper column names to match toptable 
fvarLabels(gset) <- make.names(fvarLabels(gset))
```



# order samples by group

group
```{r}
gsms <- "00112233445566"
sml <- c()
for (i in 1:nchar(gsms)) { sml[i] <- substr(gsms,i,i) }
# log2 transform
ex <- exprs(gset)
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
LogC <- (qx[5] > 100) ||
    (qx[6]-qx[1] > 50 && qx[2] > 0) ||
    (qx[2] > 0 && qx[2] < 1 && qx[4] > 1 && qx[4] < 2)
if (LogC) { 
  ex[which(ex <= 0)] <- NaN
  exprs(gset) <- log2(ex) 
}

```

## Here we measure the significance gene expression changes
the number of genes for toptable need to be chosen

```{r}
# set up the data and proceed with analysis
sml <- paste("G", sml, sep="")    # set group names
fl <- as.factor(sml)
gset$description <- fl
design <- model.matrix(~ description + 0, gset)
colnames(design) <- levels(fl)
fit <- lmFit(gset, design)
cont.matrix <- makeContrasts(G6-G0, G1-G0, G2-G1, G3-G2, G4-G3, G5-G4, G6-G5, levels=design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2, 0.01)
tT <- topTable(fit2, adjust="fdr", sort.by="B", number=1000)
tT <- subset(tT, select=c("ID","adj.P.Val","P.Value","F","Gene.symbol","Gene.title"))
```

## Here we take some exploratory data analysis.
first check to see if their expression are on the same level

```{r}
ex <- ex[ , order(sml)]
sml <- sml[order(sml)]
fl <- as.factor(sml)
labels <- c("control","4h","8h","12h","36h","48h","72h")
palette(c("#9dffd7","#ffc56e","#f3f388","#c7ff9d","#f4dff4","#dff4e4","#dcdaa5", "#AABBCC"))
#dev.new(width=4+dim(gset)[[2]]/5, height=6)
par(mar=c(2+round(max(nchar(sampleNames(gset)))/2),4,2,1))
title <- paste ("GSE4528", '/', annotation(gset), " selected samples", sep ='')
boxplot(ex, boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=fl)
legend("topleft", labels, fill=palette(), bty="n")
```

After checking, we may find genes that is most important.
```{r}

```

### Next is the clustering.
```{r}
############## hierarchical clustering ################
dist = as.dist(1 - cov1000)
hc = hclust(dist,"ave")
plot(hc)
cutree(hc,2)
cut <- cutree(hc,5)
hc5 <- data.frame(cut,names(cut))

############## k-means ################
km = kmeans(cov1000,5,nstart = 50)
km5 <- km$cluster
km5 <- data.frame(km5,cut=names(km5),symbol=as.character(tT$Gene.symbol[1:1000]))
heatmap(exnorm[,order(km5$km5)],Rowv=NA,Colv=NA,labRow=labels)

apclus <-apcluster(cov1000,details = TRUE)
show(apclus)
heatmap(
    apclus,cov1000,margins = c(10,10),Rowv = FALSE, Colv = FALSE
)

group <- factor(hc5$cut)
plotPCA(exnorm,groups=group,groupnames=levels(group))
group2 <- factor(km5$km5)
plotPCA(exnorm,groups=group2,groupnames=levels(group2))
```
